TacticToe: Learning to Prove with Tactics

### 2018-03-23 tactictoe_jv.tex
Make some edits

\section{Search Tree}\label{sec:prelim}
The application of a tactic to an initial conjecture produces a list of goals.
Each of the created goals can in turn be the input to other tactics.
Moreover, it is possible to try multiple tactics on the same goal.
In order to keep track of the progress made so far from the initial conjecture
we want to solve, we organize goals and tactics into a graph with lists of goals
as nodes and tactics as edges (See Definition~\ref{def:stree}).

In this section, we only give a description of the graph at a given
moment of the search after a certain amount of tactic applications.
Construction of the tree is done by the MCTS algorithm in
Section~\ref{sec:proofsearch}. The MCTS algorithm is guided by prediction
algorithms presented in Section~\ref{s:prediction}.

\begin{definition}\label{def:stree}(search tree)\\
A search tree $\mathfrak{T}$ is a sextuple
$(\mathbb{T},\mathbb{G},\mathbb{A},T,G,A)$
that respects the following conditions:
\begin{itemize}
\item $\mathbb{T}$ is a set of tactics, $\mathbb{G}$ is a set of goals
 and $\mathbb{A}$ is a set of nodes representing lists of goals. All objects
 are tagged with their position in the tree (see Figure~\ref{fig:choice}).
\item $T$ is a function from $\mathbb{G}$ to $\mathcal{P}(\mathbb{T})$. It
takes a goal and return the set of tactics already applied to this goal.
\item $G$ is a function from $\mathbb{A}$ to $\mathcal{P}(\mathbb{G})$.
It takes a node and returns the list of goals of this node.
\item $A$ is a function that takes a pair $(g,t)\in \mathbb{G} \times
\mathbb{T}$ such as $t \in T(g)$ and
returns a node $A(g,t)$ such as $t(g) = G(A(g,t))$. In other words, the output
$t(g)$ is exactly the list of goals contained in $A(g,t)$.
\item It is acyclic, i.e. a node cannot be a strict descendant (see
Definition~\ref{def:desc}) of itself.
\item There is exactly one node with no parents. This root node contains
exactly one goal which is the initial goal (also called conjecture).
\end{itemize}

@@

If no explicit order is given, we assume that the sets
$\mathbb{T},\mathbb{G},\mathbb{A}$ are equipped with an
arbitrary \textit{total order}. In Figure~\ref{fig:choice}, a part of the a
search tree is depicted. Goals, nodes and tactics are represented
respectively by circles, rectangles and arrows.

@@

In the following, properties are defined in the context of a
search tree $\mathfrak{T}$. In Definition~\ref{def:solved}, we specify
different status for a goal: open, pending or solved. There, we define
what a solved goal and solved node are by mutual recursion on the number of
steps it takes to solve a goal and a node.
\begin{definition}\label{def:solved}(solved goal, solved nodes, open goal,
pending goal)\\
The set of \textit{solved nodes} $\mathbb{A}^*$ and
the set of \textit{solved goals} $\mathbb{G}^*$ are defined inductively by:

\begin{align*}
\mathbb{A}^{0} &=_{def}
\lbrace a \in \mathbb{A}\ |\ G(a) = \emptyset \rbrace \\
\mathbb{G}^{0} &=_{def} \lbrace g \in \mathbb{G}\ |\
\exists t \in T(g).\ A(g,t) \in \mathbb{A}^{0} \rbrace\\
\mathbb{A}^{n+1} &=_{def} \lbrace a \in \mathbb{A}\ |\
\forall g \in G(a).\ g \in \mathbb{G}^{n} \rbrace\\
\mathbb{G}^{n+1} &=_{def} \lbrace g \in \mathbb{G}\ |\
\exists t \in T(g).\ A(g,t) \in \mathbb{A}^{n+1} \rbrace \\
\mathbb{A}^* &=_{def} \bigcup_{n \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \
\mathbb{G}^* =_{def} \bigcup_{n \in \mathbb{N}} \mathbb{G}^n\\
\end{align*}

@@

The \textit{descendants of a node} $a$ by:
\begin{align*}
\mathit{Descendant}^{0}(a) &=_{def} \lbrace a \rbrace \\
\mathit{Descendant}^{n+1}(a) &=_{def} \bigcup_{a' \in \mathit{Descendant}^{n}(a)}
\mathit{Children}(a') \\
\mathit{Descendant}(a) &=_{def} \bigcup_{n \in \mathbb{N}} \mathit{Descendant}^n(a)\\
\end{align*}

And the \textit{ancestors of a node} $a$ by:
\[\mathit{Ancestors}(a) =_{def} \lbrace b \in \mathbb{A} \ | \ a \in
Descendant(b) \rbrace\]
\end{definition}

@@

\begin{definition}\label{def:value}(Value)\\
The value of the $i^{th}$ child $a_i$ of the open goal of a parent node $p$
is determined by:
\[\mathit{Value}(a_i) = \mathit{CurEvaluation}(a_i) + c_{exploration} *
\mathit{Exploration}(a_i)\]

@@

\[\mathit{CurEvaluation}(a_i) =
  \sum_{a' \in Descendant(a_i)} \frac{\mathit{PriorEvaluation}(a')} {card\
  \mathit{Descendant}(a_i) + \mathit{Failure}(a_i)}\]

@@

\begin{align*}
D_0 &=_{def} \lbrace (t_0,g_0) \rbrace \\
D_{n+1} &=_{def} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists
(t',g') \in D_n.\ g \in t'(g') \rbrace  \\
D^* &=_{def} \bigcup_{i \in \mathbb{N}} D_i\\
\end{align*}
\end{definition}

@@

\begin{definition} (Coverage)\\
Let $\mathfrak{T}$ be the database of tactic-goal pairs. We define the
coverage $\textit{\mbox{Coverage}}(t)$ of a tactic $t$ by the number of times
this tactic
appears in
$\mathfrak{T}$. Expressing this in a formula, we get:
  \[\textit{\mbox{Coverage}}(t) =_{def} card\ \lbrace g\ |\ (t,g) \in
  \mathfrak{T} \rbrace \]
Intuitively, this notion estimates how general a tactic is by counting the
number of different goals it is useful for.
\end{definition}

\begin{definition} (Goal subsumption)\\
A goal subsumption $\le$ is a partial relation on goals.
Assuming we have a way to estimate the number of steps required to solve a goal,
a % possibly
general and useful subsumption definition is for example:
% could be:
\[g_1 \le g_2  \Leftrightarrow_{def} g_1 \mbox{ has a proof with a length
shorter
or equal to the proof of }
g_2\]
By default, we however choose for efficiency reasons a minimal subsumption defined by:
\[g_1 \le g_2  \Leftrightarrow_{def} g_1 \mbox{ is }\alpha\mbox{-equivalent
to } g_2\]
\end{definition}

We can naturally extend any goal subsumption to a subsumption on
lists of goals.
\begin{definition} (Goal list subsumption)\\
A goal list subsumption is a partial relation on lists of goals.
Given two lists of goals $l_1$ and $l_2$, we define it from the goal
subsumption $\le$ by:
%\[l_1 \le l_2  \Leftrightarrow_{def} \forall g_2 \in l_2.\ \exists g_1 \in l_1.\
%g_1 \le g_2\]
\[l_1 \le l_2  \Leftrightarrow_{def} \forall g_1 \in l_1.\ \exists g_2 \in l_2.\
g_1 \le g_2\]
\end{definition}

@@

 \[t_1 \le_g t_2 \Leftrightarrow_{def} t_1(g) \le t_2(g)\]

@@

\begin{align*}
P_{goal}(g) &=_{def} T_{sol}(g)\ \texttt{THENL}\ P_{node}(A(g,T_{sol}(g)))\\
P_{node}(a) &=_{def} [P_{goal}(g_1),\ldots,P_{goal}(g_n)]\ \ \ \text{with}\
G(a) = g_1,\ldots,g_n\\
\end{align*}

>>>>

\section{Search Tree}\label{sec:prelim}
The application of a tactic to an initial conjecture produces a list of goals.
Each of the created goals can in turn be the input to other tactics.
Moreover, it is possible to try multiple tactics on the same goal.
In order to keep track of progress made from the initial conjecture, we organize goals and tactics into a graph with lists of goals
as nodes and tactics as edges (See Definition~\ref{def:stree}).

In this section, we only give a description of the graph at a given
moment of the search, after some number of tactic applications.
Construction of the tree is done by the MCTS algorithm in
Section~\ref{sec:proofsearch}. The MCTS algorithm is guided by prediction
algorithms presented in Section~\ref{s:prediction}.

\begin{definition}\label{def:stree}(search tree)\\
A search tree $\mathfrak{T}$ is a sextuple
$(\mathbb{T},\mathbb{G},\mathbb{A},T,G,A)$
that respects the following conditions:
\begin{itemize}
\item $\mathbb{T}$ is a set of tactics, $\mathbb{G}$ is a set of goals
 and $\mathbb{A}$ is a set of nodes representing lists of goals. All objects
 are tagged with their position in the tree (see Figure~\ref{fig:choice}).
\item $T$ is a function from $\mathbb{G}$ to $\mathcal{P}(\mathbb{T})$. It
takes a goal and return the set of tactics already applied to this goal.
\item $G$ is a function from $\mathbb{A}$ to $\mathcal{P}(\mathbb{G})$.
It takes a node and returns the list of goals of this node.
\item $A$ is a function that takes a pair $(g,t)\in \mathbb{G} \times
\mathbb{T}$ such as $t \in T(g)$ and
returns a node $A(g,t)$ such that $t(g) = G(A(g,t))$. In other words, the output
$t(g)$ is exactly the list of goals contained in $A(g,t)$.
\item It is acyclic, i.e., a node cannot be a strict descendant (see
Definition~\ref{def:desc}) of itself.
\item There is exactly one node with no parents. This root node contains
exactly one goal which is the initial goal (conjecture).
\end{itemize}

@@

If no explicit order is given, we assume that the sets
$\mathbb{T},\mathbb{G},\mathbb{A}$ are equipped with an
arbitrary \textit{total order}. Figure~\ref{fig:choice} depicts part of a
search tree. Goals, nodes and tactics are represented
respectively by circles, rectangles and arrows.

@@

In the following, properties are defined in the context of a
search tree $\mathfrak{T}$. In Definition~\ref{def:solved}, we specify
different states for a goal: open, pending or solved. There, we define
what a solved goal and solved node are by mutual recursion on the number of
steps it takes to solve a goal and a node.
\begin{definition}\label{def:solved}(solved goal, solved nodes, open goal,
pending goal)\\
The set of \textit{solved nodes} $\mathbb{A}^*$ and
the set of \textit{solved goals} $\mathbb{G}^*$ are defined inductively by:

\begin{align*}
\mathbb{A}^{0} &=_{\mathrm{def}}
\lbrace a \in \mathbb{A}\ |\ G(a) = \emptyset \rbrace \\
\mathbb{G}^{0} &=_{\mathrm{def}} \lbrace g \in \mathbb{G}\ |\
\exists t \in T(g).\ A(g,t) \in \mathbb{A}^{0} \rbrace\\
\mathbb{A}^{n+1} &=_{\mathrm{def}} \lbrace a \in \mathbb{A}\ |\
\forall g \in G(a).\ g \in \mathbb{G}^{n} \rbrace\\
\mathbb{G}^{n+1} &=_{\mathrm{def}} \lbrace g \in \mathbb{G}\ |\
\exists t \in T(g).\ A(g,t) \in \mathbb{A}^{n+1} \rbrace \\
\mathbb{A}^* &=_{\mathrm{def}} \bigcup_{n \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \
\mathbb{G}^* =_{\mathrm{def}} \bigcup_{n \in \mathbb{N}} \mathbb{G}^n\\
\end{align*}

@@

The \textit{descendants of a node} $a$ by:
\begin{align*}
\mathit{Descendant}^{0}(a) &=_{\mathrm{def}} \lbrace a \rbrace \\
\mathit{Descendant}^{n+1}(a) &=_{\mathrm{def}} \bigcup_{a' \in \mathit{Descendant}^{n}(a)}
\mathit{Children}(a') \\
\mathit{Descendant}(a) &=_{\mathrm{def}} \bigcup_{n \in \mathbb{N}} \mathit{Descendant}^n(a)\\
\end{align*}

And the \textit{ancestors of a node} $a$ by:
\[\mathit{Ancestors}(a) =_{\mathrm{def}} \lbrace b \in \mathbb{A} \ | \ a \in
Descendant(b) \rbrace\]
\end{definition}

@@

\begin{definition}\label{def:value}(Value)\\
The value of the $i^{th}$ child $a_i$ of the open goal of a parent node $p$
is determined by:
  \[\mathit{Value}(a_i) = \mathit{CurEvaluation}(a_i) + c_{\mathit{exploration}} *
\mathit{Exploration}(a_i)\]

@@

\[\mathit{CurEvaluation}(a_i) =
  \sum_{a' \in \mathit{Descendant}(a_i)} \frac{\mathit{PriorEvaluation}(a')} {card\
  \mathit{Descendant}(a_i) + \mathit{Failure}(a_i)}\]

@@

\begin{align*}
D_0 &=_{\mathrm{def}} \lbrace (t_0,g_0) \rbrace \\
D_{n+1} &=_{\mathrm{def}} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists
(t',g') \in D_n.\ g \in t'(g') \rbrace  \\
D^* &=_{\mathrm{def}} \bigcup_{i \in \mathbb{N}} D_i\\
\end{align*}
\end{definition}

@@

\begin{definition} (Coverage)\\
Let $\mathfrak{T}$ be the database of tactic-goal pairs. We define the
coverage $\textit{\mbox{Coverage}}(t)$ of a tactic $t$ by the number of times
this tactic
appears in
$\mathfrak{T}$. Expressing this in a formula, we get:
\[\textit{\mbox{Coverage}}(t) =_{\mathrm{def}} card\ \lbrace g\ |\ (t,g) \in
  \mathfrak{T} \rbrace \]
Intuitively, this notion estimates how general a tactic is by counting the
number of different goals it is useful for.
\end{definition}

\begin{definition} (Goal subsumption)\\
A goal subsumption $\le$ is a partial relation on goals.
Assuming we have a way to estimate the number of steps required to solve a goal,
a % possibly
general and useful subsumption definition is for example:
% could be:
\[g_1 \le g_2  \Leftrightarrow_{\mathrm{def}} g_1 \mbox{ has a proof with a length
shorter
or equal to the proof of }
g_2\]
By default, we however choose for efficiency reasons a minimal subsumption defined by:
\[g_1 \le g_2  \Leftrightarrow_{\mathrm{def}} g_1 \mbox{ is }\alpha\mbox{-equivalent
to } g_2\]
\end{definition}

We can naturally extend any goal subsumption to a subsumption on
lists of goals.
\begin{definition} (Goal list subsumption)\\
A goal list subsumption is a partial relation on lists of goals.
Given two lists of goals $l_1$ and $l_2$, we define it from the goal
subsumption $\le$ by:
%\[l_1 \le l_2  \Leftrightarrow_{\mathrm{def}} \forall g_2 \in l_2.\ \exists g_1 \in l_1.\
%g_1 \le g_2\]
\[l_1 \le l_2  \Leftrightarrow_{\mathrm{def}} \forall g_1 \in l_1.\ \exists g_2 \in l_2.\
g_1 \le g_2\]
\end{definition}

@@

\[t_1 \le_g t_2 \Leftrightarrow_{\mathrm{def}} t_1(g) \le t_2(g)\]

@@

\begin{align*}
P_{goal}(g) &=_{\mathrm{def}} T_{sol}(g)\ \texttt{THENL}\ P_{node}(A(g,T_{sol}(g)))\\
P_{node}(a) &=_{\mathrm{def}} [P_{goal}(g_1),\ldots,P_{goal}(g_n)]\ \ \ \text{with}\
G(a) = g_1,\ldots,g_n\\
\end{align*}
